---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# havel

<!-- badges: start -->
<!-- badges: end -->

The goal of havel is to vizualize and tabulate dependencies in R packages. This will help R package developers identify dependencies that contribute more to the dependency load and/or are easier to remove.

## Installation

You can install the development version of havel from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("andrewGhazi/havel")
```

## Example

Use `plot_deps_graph()` to look up and plot the dependency graph of a package:

```{r, echo=FALSE}
set.seed(1)
```

```{r example, out.width="800px", dpi=300, message=FALSE}
library(havel)

plot_deps_graph("ggplot2")
```

You can see on the left side that the `scales` package contributes several dependencies that nothing else depends on. If the `ggplot2` developers were looking to lighten their package (not that they should), eliminating the dependency on `scales` might be a good place to start looking.

< bit about `uniq_pkg_deps()` >

## Related/prior work

There are other packages / posts out there on this type of thing:

* [`miniCRAN`](https://github.com/andrie/miniCRAN) has a [vignette on similar plotting functionality](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-dependency-graph.html).
* [`pkgnet`](https://uptake.github.io/pkgnet/index.html) offers an expansive dependency analysis suite.
* [`depgraph`](https://github.com/crsh/depgraph) is pretty similar but uses different aesthetic mappings.
* [`DependenciesGraphs`](https://datastorm-open.github.io/DependenciesGraphs/) and [`pkgdepR`](https://cran.r-project.org/web/packages/pkgdepR/index.html) visualize the dependencies of *functions* used within a package.
* [`pkgndep`](https://cran.r-project.org/web/packages/pkgndep/index.html) visualizes similar information in a heatmap format, as well as provides different "heaviness" metrics.

`havel` offers some unique advantages nonetheless:

* It uses `pak::pkg_deps()` to look up the dependency connections instead of `tools::package_dependencies()`. This enables more versatile lookups of packages on Bioc, GitHub, etc.
* `havel` itself is fairly light and fast, depending only on pak, some fastverse packages, and base R graphics...
  * ...but it includes ggplot2 in the Suggests if you prefer that. Set `gg = TRUE`.
* The default plotting parameters are carefully hand-crafted to look nice 😊

## Misc

> What's with the name?

Havel's Ring is an item in Dark Souls that helps you deal with excessive burdens.

## TODO

* more tests
* color edges by dependency type
* copy over ggplot version
* uniq_pkg_deps() function
